\begin{thebibliography}{99}


	\bibitem{ot3manual}
		Argent Data Systems,
		\emph{Tracker3 Family User's Manual}.
		Santa Maria, California,
		Revised 1-3-2014.
		\url{http://www.argentdata.com/support/tracker3_manual.pdf}.

	\bibitem{bobhatesnice}
		R. Bruninga,
		``Beacon rate feedback."
		Online posting. 24 Nov. 2008. APRSsig.
		\url{http://www.tapr.org/pipermail/aprssig/2008-November/027595.html}.

	\bibitem{millernopid}
		S. Miller,
		``AX25 Protocol Group."
		Online posting. 2 Jun. 2008. APRSsig.
		\url{http://www.tapr.org/pipermail/aprssig/2008-June/025220.html}.

	\bibitem{millerinterview}
		S. Miller. Personal interview. 25 Mar. 2014.

	\bibitem{aprsspec}
		I. Wade, Ed,
		\emph{Automatic Position Reporting System --- 
		APRS Protocol Reference Protocol Version 1.0}.
		Approved Version 1.0.1.
		Tucson, Arizona: Tucson Amateur Packet Radio Corp, 2000. 
		\url{http://www.aprs.org/doc/APRS101.PDF}

	\bibitem{ax25spec}
		W. Beech, et al.,
		\emph{AX.25 Link Access Protocol for Amateur Packet Radio Version 2.2}.
		Tucson, Arizona: Tucson Amateur Packet Radio Corp, 1998. 
		\url{http://www.tapr.org/pdf/AX25.2.2.pdf}

	\bibitem{packetthroughput}
		N. Abramson,
		``The Throughput of Packet Broadcasting Channels," in
		\emph{IEEE Transactions on Communications,}
		Vol. COM-25, no. 1, Jan. 1977, pp.~117-128.

	\bibitem{failureofpoisson}
		V. Paxson, S. Floyd,
		``Wide-Area Traffic: The Failure of Poisson Modeling," in
		\emph{IEEE Transactions on Networking,}
		3(3), Jun. 1995, pp.~226-244.

	\bibitem{smartbeacon}
		S. Bragg, 
		``SmartBeaconing\texttrademark,"
		4 Aug. 2007.
		\url{http://www.hamhud.net/hh2/smartbeacon.html}

	\bibitem{smartbeaconwiki}
		C. Mills, J. Ewen,
		``SmartBeaconing,"
		5 Jan. 2012.
		\url{http://info.aprs.net/index.php?title=SmartBeaconing}

	\bibitem{smartbeaconemail1}
		S. Bragg, ``Question on SmartBeacon Algorithm for Thesis."
		Private email. 3 May 2014.

	\bibitem{nmearmc}
		D. DePriest,
		``NMEA data."
		\url{http://www.gpsinformation.org/dale/nmea.htm#RMC}

	\bibitem{KISSspec}
		M. Chepponis, P. Karn,
		``The KISS TCN: A simple Host-to-TNC communications protocol,"
		in \emph{ARRL 6th Computer Networking Conference},
		1987, pp.~38-43.
		Translated to HTML Jan. 1997 by P. Karn,
		\url{http://ax25.net/kiss.aspx}

	\bibitem{TNC2manual}
		Tucson Amateur Packet Radio Corporation,
		\emph{Terminal Node Controller TNC2 Assembly Manual}.
		Tucson, Arizona,
		June 1991.

	\bibitem{iso13239}
		\emph{Information technoloy --- Telecommunications and information
			exchange between systems --- High-level data link control (HDCL)
		procedures}, ISO Standard 13239, 2002.

	\bibitem{202tspec}
		American Telephone and Telegraph Company,
		\emph{Data Sets 202S and 202T Interface Specification},
		AT\&T Publication 41212,
		July, 1976.

	\bibitem{n1vgphy}
		S. Miller,
		``1200 Baud Packet Radio Details."
		\url{http://n1vg.net/packet/index.php}

	\bibitem{aprsunveiled}
		B. Simmons,
		``APRS Unveiled," in \emph{QEX},
		pp.~19-23,
		Nov./Dec. 2012.

	\bibitem{ituv42}
		\emph{Error-correcting procedures for DCEs using 
		asynchronous-to-synchronous converion}, ITU-T standard V.42, 2002.

\end{thebibliography}

