\chapter{TNC2 Monitor}

The TNC2 was a terminal node controller developed by TAPR which
quickly became the definitive TNC in the amateur packet network.
This popularity meant that it enjoyed becoming the de-facto 
standard for many of the commands and interfaces involving 
TNCs.

One of the de-facto standards solidified by the TNC2 is the text
representation of AX.25 packets, which the TNC2 used in what it
called ``monitor mode." This mode asked that the TNC print 
every packet it received via RF to the local console such that
users can see and monitor the other AX.25 traffic on the channel.
While the TNC2 monitor mode does not capture the entire state of 
an AX.25 header, APRS uses so little of the header (as discussed in
the AX.25 chapter) that the TNC2 markup proved sufficient as the 
format used to pass APRS packets throughout the APRS-Internet System
backbone created to aggregate APRS traffic online.

Each packet must consist of a source address, a destination address,
an optional list of digipeaters, and the packet payload.
The source and destination calls are separated by a right chevron,
followed by a comma separated list of digital repeaters, and finally
terminated by a colon before the higher layer payload.

The digipeaters are listed in sequential order, and an asterisk
is appended to at least the last digipeater in the 
routing path that has already processed and re-transmitted the packet.
Some implementations of the TNC2 monitor mode append the ``consumed"
asterisk to every used digipeater in addition to the last one,
which semantically makes no difference due to digipeaters
being consumed in a strictly monotonic order.

TODO I think the APRS-IS uses the later format.

SRCCALL\textgreater{}DSTCALL,DIGI*,DIGI:,Packet payload
